---
title: "Data Cleaning"
author: "Sean F. Larsen"
date: "February 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("readxl")
library("dplyr")
library("tidyr")
```
# The Human Trafficking Project

Human trafficking is the business of stealing freedom for profit. In some cases, traffickers trick, defraud or physically force victims into providing commercial sex. In others, victims are lied to, assaulted, threatened or manipulated into working under inhumane, illegal or otherwise unacceptable conditions. It is a multi-billion dollar criminal industry that denies freedom to 24.9 million people around the world.

My project is to examine the data avaialable on the FBI website for human trafficking.

## Data Cleaning

I will be using the readxl, dplyr, and tdyr packages for the data cleaning process.

Human Trafficking files were loaded from the FBI Uniform Crime Report Crimes in the United States data site.
Each year has it's own page and explanation of the data and methods of gathering.
Information and datacan be accessed from the following sites:

<https://ucr.fbi.gov/crime-in-the-u.s/2014/crime-in-the-u.s.-2014/additional-reports/human-trafficking-report/human-trafficking.pdf>

<https://ucr.fbi.gov/crime-in-the-u.s/2015/crime-in-the-u.s.-2015/additional-reports/human-trafficking/humantrafficking_-2015-_final>

<https://ucr.fbi.gov/crime-in-the-u.s/2016/crime-in-the-u.s.-2016/additional-publications/human-trafficking>

<https://ucr.fbi.gov/crime-in-the-u.s/2017/crime-in-the-u.s.-2017/additional-data-collections/human-trafficking>

2018 data is not available at this time

The follwoing Tables are the Offense and Clearnace using the FBI's Summary Reporting System (SRS)
When more than one crime was committed by the same person or group of persons.  The time and space intervals
separating the crimes were insignificant.  So, the crime highest in the hierarchy is the only offense reported.

An Offense is reported and under investigation by a participating organization.
A Clearance is an offense where a perpetrator has been arrested or convicted.

## Read and Write

This code reads each file and creates a data frame to work with.
```{r include=TRUE}
ht_offenses_clearing_2014 <- read_excel("Table_2_Human_Trafficking_Offenses_and_Clearances_by_State_2014.xlsx")
ht_offenses_clearing_2014_df <- data.frame(ht_offenses_clearing_2014)
write.csv(ht_offenses_clearing_2014, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2014.csv")

ht_offenses_clearing_2015 <- read_excel("table-1_Human_Trafficking_Offenses_and_Clearances_by_State_2015.xls")
ht_offenses_clearing_2015_df <- data.frame(ht_offenses_clearing_2015)
write.csv(ht_offenses_clearing_2015, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2015.csv")

ht_offenses_clearing_2016 <- read_excel("table-1_Human_Trafficking_Offenses_and_Clearances_by_State_2016.xls")
ht_offenses_clearing_2016_df <- data.frame(ht_offenses_clearing_2016)
write.csv(ht_offenses_clearing_2016, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2016.csv")

ht_offenses_clearing_2017 <- read_excel("table-1_Human_Trafficking_Offenses_and_Clearances_by_State_2017.xls")
ht_offenses_clearing_2017_df <- data.frame(ht_offenses_clearing_2017)
write.csv(ht_offenses_clearing_2017, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2017.csv")
```

## Column Names

There are no column names, so I am adding the column names to each file.

```{r include=TRUE}
colnames(ht_offenses_clearing_2014_df) <- c("state", "commercial_sex_act_offense", "commercial_sex_act_cleared", "commercial_sex_act_cleared_under_18", "involuntary_servitude_offense", "involuntary_servitude_cleared", "involuntary_servitude_cleared_under_18","total_offenses", "total_cleared","total_under_18")

colnames(ht_offenses_clearing_2015_df) <- c("state", "commercial_sex_act_offense", "commercial_sex_act_cleared", "commercial_sex_act_cleared_under_18", "involuntary_servitude_offense", "involuntary_servitude_cleared", "involuntary_servitude_cleared_under_18","total_offenses", "total_cleared","total_under_18")

colnames(ht_offenses_clearing_2016_df) <- c("state", "commercial_sex_act_offense", "commercial_sex_act_cleared", "commercial_sex_act_cleared_under_18", "involuntary_servitude_offense", "involuntary_servitude_cleared", "involuntary_servitude_cleared_under_18","total_offenses", "total_cleared","total_under_18")

colnames(ht_offenses_clearing_2017_df) <- c("state", "commercial_sex_act_offense", "commercial_sex_act_cleared", "commercial_sex_act_cleared_under_18", "involuntary_servitude_offense", "involuntary_servitude_cleared", "involuntary_servitude_cleared_under_18","total_offenses", "total_cleared","total_under_18")
```

## Rows that aren't needed

There are 4 rows in each file that were used a data information header in excel, that are not needed.  I am deleting these rows using the following code.

```{r include=TRUE}
ht_offenses_clearing_2014_df <- ht_offenses_clearing_2014_df[-c(1, 2, 3, 4),]
ht_offenses_clearing_2015_df <- ht_offenses_clearing_2015_df[-c(1, 2, 3, 4),]
ht_offenses_clearing_2016_df <- ht_offenses_clearing_2016_df[-c(1, 2, 3, 4),]
ht_offenses_clearing_2017_df <- ht_offenses_clearing_2017_df[-c(1, 2, 3, 4),]
```

## Adding Columns

Since each file represents data form a particular year, and eventually the data will be combined, I need a column for the year.
This is the code I am using to create the new column and add the year of the file as the data in the new column.

```{r include=TRUE}
# Creating a new column for the year and move that column to the front.
ht_offenses_clearing_2014_df <- mutate(ht_offenses_clearing_2014_df,"year" = 2014)
ht_offenses_clearing_2014_df <- select(ht_offenses_clearing_2014_df, year, everything())

ht_offenses_clearing_2015_df <- mutate(ht_offenses_clearing_2015_df,"year" = 2015)
ht_offenses_clearing_2015_df <- select(ht_offenses_clearing_2015_df, year, everything())

ht_offenses_clearing_2016_df <- mutate(ht_offenses_clearing_2016_df,"year" = 2016)
ht_offenses_clearing_2016_df <- select(ht_offenses_clearing_2016_df, year, everything())

ht_offenses_clearing_2017_df <- mutate(ht_offenses_clearing_2017_df,"year" = 2017)
ht_offenses_clearing_2017_df <- select(ht_offenses_clearing_2017_df, year, everything())
```

## Writing the data to a file

Now the data hass been set-up, I will rewrite each of the CSV files.

```{r include=FALSE}
write.csv(ht_offenses_clearing_2014, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2014.csv")
write.csv(ht_offenses_clearing_2015, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2015.csv")
write.csv(ht_offenses_clearing_2016, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2016.csv")
write.csv(ht_offenses_clearing_2017, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2017.csv")
```

# Combine and Sort

## This code combines all four data frames into a single dataframe to use with analysis, and writes it to a single CSV file.

```{r include=TRUE}
ht_offenses_clearing_4_years_df <- bind_rows(ht_offenses_clearing_2014_df, ht_offenses_clearing_2015_df, ht_offenses_clearing_2016_df, ht_offenses_clearing_2017_df)
write.csv(ht_offenses_clearing_4_years_df, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2014_to_2017.csv")
```

## How the Data is Organized

Now that the data is organized and combined and after reading the FBI website on the process of compiling the data,
I've leanred that data is generally broken down by age and is categorized as adult and juvenile.
If an individual is over 18 they are adults and under 18 they are juveniles.
There are also two tpyes of offences listed.  Commercial Sex Acts and Involuntary Servitude.
Both age and offense need to be separated by columns.

## Separate Offenses and Age

I am going to pull out Commercial Sex Offense and Involuntary servituded into two different groups and then combine them back.

```{r include=TRUE}
comm_sex_offense_df <- select(ht_offenses_clearing_4_years_df, "year", "state", "commercial_sex_act_offense", "commercial_sex_act_cleared", "commercial_sex_act_cleared_under_18")
```

This code removes the invalid row at the bottom
```{r include=TRUE}
comm_sex_offense_df <- comm_sex_offense_df[-c(149),]
```
This code creates a new column for offense
```{r include=TRUE}
comm_sex_offense_df <- mutate(comm_sex_offense_df, "offense" = "commercial sex act")
```

I discovered that my values are characters and not numeric.  This code Changes the columns to numeric in value.

```{r include=TRUE}
comm_sex_offense_df$year <- as.numeric(comm_sex_offense_df$year)
comm_sex_offense_df$commercial_sex_act_offense <- as.numeric(comm_sex_offense_df$commercial_sex_act_offense)
comm_sex_offense_df$commercial_sex_act_cleared <- as.numeric(comm_sex_offense_df$commercial_sex_act_cleared)
comm_sex_offense_df$commercial_sex_act_cleared_under_18 <- as.numeric(comm_sex_offense_df$commercial_sex_act_cleared_under_18)
```

This code creates a new column for adult cleared offenses, then reorders the columns

```{r include=TRUE}
comm_sex_offense_df <- mutate(comm_sex_offense_df, commercial_sex_act_cleared_adults = (commercial_sex_act_cleared - commercial_sex_act_cleared_under_18))

comm_sex_offense_df <- select(comm_sex_offense_df, year, state, offense,commercial_sex_act_offense, commercial_sex_act_cleared, commercial_sex_act_cleared_adults, commercial_sex_act_cleared_under_18, everything())
```

With commercial_sex_act_cleared_adults separted out from commercial_sex_act_cleared, commercial_sex_act_cleared is no longer needed.
The columns also need to be rename to help understanding the data in each column.

```{r include=TRUE}
comm_sex_offense_df$commercial_sex_act_cleared <- NULL
colnames(comm_sex_offense_df) <- c("year", "state", "offense", "occurrence", "adult", "juvenile")
```

This will create and column for age and gather Adult and Juvenile.

```{r include=TRUE}
comm_sex_offense_df <- gather(comm_sex_offense_df, age, cleared, adult, juvenile)
```

Now I'm pulling the data for involuntary servitude.  I will clean this up and then put coomercial sex act and involuntary servitude back together.

```{r include=TRUE}
invol_serv_df <- select(ht_offenses_clearing_4_years_df, "year", "state", "involuntary_servitude_offense", "involuntary_servitude_cleared", "involuntary_servitude_cleared_under_18")
```

This code removes the invalid row at the bottom.

```{r include=TRUE}
invol_serv_df <- invol_serv_df[-c(149),]
```

This code creates a new column for offenses.

```{r include=TRUE}
invol_serv_df <- mutate(invol_serv_df,"offense" = "involuntary servitude")
```

This code Changes certain columns to numeric values.

```{r include=TRUE}
invol_serv_df$year <- as.numeric(invol_serv_df$year)
invol_serv_df$involuntary_servitude_offense <- as.numeric(invol_serv_df$involuntary_servitude_offense)
invol_serv_df$involuntary_servitude_cleared <- as.numeric(invol_serv_df$involuntary_servitude_cleared)
invol_serv_df$involuntary_servitude_cleared_under_18 <- as.numeric(invol_serv_df$involuntary_servitude_cleared_under_18)
```

this code creates a new column for adult cleared offenses.

```{r include=TRUE}
invol_serv_df <- mutate(invol_serv_df, involuntary_servitude_cleared_adult = (involuntary_servitude_cleared - involuntary_servitude_cleared_under_18))
```

With involuntary_servitude_cleared_adult separted out from involuntary_servitude_cleared, involuntary_servitude_cleared is no longer needed. This code removes the column, reoders the columns, and renames them for easier analysis later.

```{r include=TRUE}
invol_serv_df$involuntary_servitude_cleared <- NULL
invol_serv_df <- select(invol_serv_df, year, state, offense, involuntary_servitude_offense, involuntary_servitude_cleared_adult, involuntary_servitude_cleared_under_18, everything())
colnames(invol_serv_df) <- c("year", "state", "offense", "occurrence", "adult", "juvenile")
```

This will create and column for age and gather Adult and Juvenile.
```{r include=TRUE}
invol_serv_df <- gather(invol_serv_df, age, cleared, adult, juvenile)
```

# Combine

Now this code combines the two data frames into one final data frame.

```{r include=TRUE}
ht_offenses_4_years_clean_df <- bind_rows(comm_sex_offense_df, invol_serv_df)
```

# One More Look

After examining the data, I noticed a few states have an asterisk after it in certain years.
This code removes the astreisk.

```{r include=TRUE}
ht_offenses_4_years_clean_df$state <- gsub("U. S. Virgin Islands", "U.S. Virgin Islands", ht_offenses_4_years_clean_df$state)
```

U.S. Virgin Islands is spelled two Different Ways.  This will correct the spelling of the odd one.

```{r include=TRUE}
ht_offenses_4_years_clean_df$state <- gsub("\\*", "", ht_offenses_4_years_clean_df$state)
```

# FBI Regions

Currently the data is all listed by state. For possible simplification of analysis, this code addes the twelve FBI regions to the data.  The FBI uses a number code, however I will add an additional column for a description. Then the code will reoder the columns.

```{r include=TRUE}
ht_offenses_4_years_clean_df <- mutate(ht_offenses_4_years_clean_df, region = state)
ht_offenses_4_years_clean_df <- mutate(ht_offenses_4_years_clean_df, region_description = state)
```

This code will replace the State with the Region code and the region description.

```{r include=TRUE}
ht_offenses_4_years_clean_df$region <- gsub("Maine", "1", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Maine", "New England", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("New Hampshire", "1", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("New Hampshire", "New England", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Rhode Island", "1", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Rhode Island", "New England", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Massachusetts", "1", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Massachusetts", "New England", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("New York", "2", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("New York", "Mideast North", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Vermont", "2", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Vermont", "Mideast North", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Connecticut", "2", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Connecticut", "Mideast North", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Pennsylvania", "3", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Pennsylvania", "Mideast South", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("New Jersey", "3", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("New Jersey", "Mideast South", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Delaware", "3", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Delaware", "Mideast South", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("West Virginia", "4", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("West Virginia", "Mid-Atlantic", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Maryland", "4", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Maryland", "Mid-Atlantic", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Virginia", "4", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Virginia", "Mid-Atlantic", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("North Carolina", "4", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("North Carolina", "Mid-Atlantic", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("South Carolina", "4", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("South Carolina", "Mid-Atlantic", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Alabama", "5", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Alabama", "Gulf Coast East", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Mississippi", "5", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Mississippi", "Gulf Coast East", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Georgia", "5", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Georgia", "Gulf Coast East", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Florida", "5", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Florida", "Gulf Coast East", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Puerto Rico", "5", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Puerto Rico", "Gulf Coast East", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("U.S. Virgin Islands", "5", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("U.S. Virgin Islands", "Gulf Coast East", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Michigan", "6", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Michigan", "Central", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Ohio", "6", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Ohio", "Central", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Kentucky", "6", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Kentucky", "Central", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Tennessee", "6", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Tennessee", "Central", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Wisconsin", "7", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Wisconsin", "North Central", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Indiana", "7", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Indiana", "North Central", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Illinois", "7", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Illinois", "North Central", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Minnesota", "7", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Minnesota", "North Central", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("North Dakota", "8", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("North Dakota", "Plains", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("South Dakota", "8", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("South Dakota", "Plains", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Nebraska", "8", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Nebraska", "Plains", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Iowa", "8", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Iowa", "Plains", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Missouri", "8", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Missouri", "Plains", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Arkansas", "8", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Arkansas", "Plains", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("California", "9", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("California", "Southwest/Pacific", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Nevada", "9", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Nevada", "Southwest/Pacific", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Arizona", "9", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Arizona", "Southwest/Pacific", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Hawaii", "9", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Hawaii", "Southwest/Pacific", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Alaska", "9", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Alaska", "Southwest/Pacific", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Guam", "9", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Guam", "Southwest/Pacific", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Wyoming", "10", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Wyoming", "Rockies/Mid-West", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Utah", "10", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Utah", "Rockies/Mid-West", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Colorado", "10", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Colorado", "Rockies/Mid-West", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Kansas", "10", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Kansas", "Rockies/Mid-West", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("New Mexico", "10", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("New Mexico", "Rockies/Mid-West", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Oklahoma", "10", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Oklahoma", "Rockies/Mid-West", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Texas", "11", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Texas", "Gulf Coast West", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Louisiana", "11", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Louisiana", "Gulf Coast West", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Washington", "12", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Washington", "Northeast", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Oregon", "12", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Oregon", "Northeast", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Idaho", "12", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Idaho", "Northeast", ht_offenses_4_years_clean_df$region_description)
ht_offenses_4_years_clean_df$region <- gsub("Montana", "12", ht_offenses_4_years_clean_df$region)
ht_offenses_4_years_clean_df$region_description <- gsub("Montana", "Northeast", ht_offenses_4_years_clean_df$region_description)
```
This code reorders the column names for later analysis.

```{r include=TRUE}
ht_offenses_4_years_clean_df <- select(ht_offenses_4_years_clean_df, year, region, region_description, state, offense, occurrence, age, cleared,everything())
```

# Write the CSV file

This code writes the final data to a new CSV file.
```{r include=TRUE}
write.csv(ht_offenses_4_years_clean_df, file = "Human_Trafficking_Offenses_and_Clearances_by_State_2014_to_2017_clean.csv")
```
