---
title: "Human Trafficking Data Analysis"
author: "Sean F. Larsen"
date: "February 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
```

## Retriving the Data

The data has been cleaned and is ready for analysis.  First step will be to look at the data and determine what additional data will be needed.

```{r include=TRUE}
ht_offenses_4_years_clean <- read.csv("Human_Trafficking_Offenses_and_Clearances_by_State_2014_to_2017_clean.csv")
ht_offenses_4_years_clean_df <- data.frame(ht_offenses_4_years_clean)
ht_offenses_4_years_clean_df$X <- NULL
```

## Crime Stats

In the United States Crime Statisitics are ussually descibed as instance per 100,000 people in a population.
This code will add a column dividing the instances of the crimes by the population and multiplying by 100,000.
This is the number that will be used in the analysis.

```{r include = TRUE}
ht_offenses_4_years_clean_df <- mutate(ht_offenses_4_years_clean_df, occ_per_100k = occurrence / population * 100000)
ht_offenses_4_years_clean_df <- mutate(ht_offenses_4_years_clean_df, clr_per_100k = cleared / population * 100000)
```

First Attempt at a Plot
